---
title: "Simple document"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Read in some data
read in the litters dataset.
```{r}
litters_df = read_csv("FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
in the janitor package, use clean_names


## take a look at the data
printing in the console
```{r}
litters_df 
``` 

```{r}
head(litters_df)
```

```{r}
tail(litters_df)
```

```{r}
skimr::skim(litters_df)
```
In the skimr package use the skim function

## OPtions to read_csv
```{r}
litters_df = read_csv("FAS_litters.csv", skip=10, col_names=FALSE)
```
check out '?read_csv()' for more information.


## Other file formats
Read in an excel file
```{r}
library(readxl)
mlb_df = read_excel("mlb11.xlsx")
mlb_df
```

read in a SAS file
```{r}
library(haven)
pulse_df = read_sas("./public_pulse_data.sas7bdat")
pulse_df
```








-------------------------------------------------
I'm an R Markdown document! 

# Section 1

Here's a **code chunk** that samples from 
a _normal distribution_:

```{r}
samp = rnorm(100)
length(samp)
```

# Section 2

I can take the mean of the sample, too!
The mean is `r mean(samp)`.