---
title: "Simple document"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Read in some data
read in the litters dataset.
```{r}
litters_df = read_csv("FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
in the janitor package, use clean_names


## take a look at the data
printing in the console
```{r}
litters_df 
``` 

```{r}
head(litters_df)
```

```{r}
tail(litters_df)
```

```{r}
skimr::skim(litters_df)
```
In the skimr package use the skim function

## OPtions to read_csv
```{r}
litters_df = read_csv("FAS_litters.csv", skip=10, col_names=FALSE)
```
check out '?read_csv()' for more information.


## Other file formats
Read in an excel file
```{r}
library(readxl)
mlb_df = read_excel("mlb11.xlsx")
mlb_df
```

read in a SAS file
```{r}
library(haven)
pulse_df = read_sas("./public_pulse_data.sas7bdat")
pulse_df
```
# comparison with base r

what about "read.csv"? just never use read.csv, always use read_csv
```{r}
litters_base = read.csv("FAS_litters.csv")
litters_readr = read_csv("FAS_litters.csv")

litters_base
Litters_readr
```

## Exporting data
export the mlb sub-table.
```{r}
write_csv(mlb_df, "./mlb_subtable.csv")
```

## Data Manipulation
## Select
choose some columns and not others.
```{r}
select(litters_df, group, gd0_weight)

# specify a range of columns use ":"
select(litters_df, group, gd0_weight:gd_of_birth)
```

only get rid of the "little_number" column where keep everything else
```{r}
select(litters_df, -litter_number)
```

Renaming columns
this one is to select and rename simultaneously
```{r}
select(litters_df, GROUP= group, LITTer_NUmber=litter_number)
```
```{r}
rename (litters_df, GROUP= group, LITTer_NUmber=litter_number )
```

only select columns that start with gd
```{r}
select(litters_df, starts_with("gd"))
```

this would select litter_number and move it from whereever
to the beginning and keep everything else
```{r}
select(litters_df, litter_number, everything())
```
just trying to relocate this "litter_number" column to the beginning
```{r}
relocate(litters_df, litter_number)
```














-------------------------------------------------
I'm an R Markdown document! 

# Section 1

Here's a **code chunk** that samples from 
a _normal distribution_:

```{r}
samp = rnorm(100)
length(samp)
```

# Section 2

I can take the mean of the sample, too!
The mean is `r mean(samp)`.