---
title: "Simple document"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Read in some data
read in the litters dataset.
```{r}
litters_df = read_csv("FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
in the janitor package, use clean_names


## take a look at the data
printing in the console
```{r}
litters_df 
``` 

```{r}
head(litters_df)
```

```{r}
tail(litters_df)
```

```{r}
skimr::skim(litters_df)
```
In the skimr package use the skim function

## OPtions to read_csv
```{r}
litters_df = read_csv("FAS_litters.csv", skip=10, col_names=FALSE)
```
check out '?read_csv()' for more information.


## Other file formats
Read in an excel file
```{r}
library(readxl)
mlb_df = read_excel("mlb11.xlsx")
mlb_df
```

read in a SAS file
```{r}
library(haven)
pulse_df = read_sas("./public_pulse_data.sas7bdat")
pulse_df
```
# comparison with base r

what about "read.csv"? just never use read.csv, always use read_csv
```{r}
litters_base = read.csv("FAS_litters.csv")
litters_readr = read_csv("FAS_litters.csv")

litters_base
Litters_readr
```

## Exporting data
export the mlb sub-table.
```{r}
write_csv(mlb_df, "./mlb_subtable.csv")
```

## Data Manipulation
## Select
choose some columns and not others.Keep the rows intact
```{r}
select(litters_df, group, gd0_weight)

# specify a range of columns use ":"
select(litters_df, group, gd0_weight:gd_of_birth)
```

only get rid of the "little_number" column where keep everything else
```{r}
select(litters_df, -litter_number)
```

Renaming columns
this one is to select and rename simultaneously
```{r}
select(litters_df, GROUP= group, LITTer_NUmber=litter_number)
```
```{r}
rename (litters_df, GROUP= group, LITTer_NUmber=litter_number )
```

only select columns that start with gd
```{r}
select(litters_df, starts_with("gd"))
```

this would select litter_number and move it from whereever
to the beginning and keep everything else
```{r}
select(litters_df, litter_number, everything())
```
just trying to relocate this "litter_number" column to the beginning
```{r}
relocate(litters_df, litter_number)
```

## filter
remove rows or keep columns
```{r}
filter(litters_df, gd0_weight < 22)
#only keep the gestation weight that's less than 22

filter(litters_df, gd0_weight >= 22)
```
```{r}
filter(litters_df, gd_of_birth==20)
#gestation date of birth is equal to 20, use "=="
```

```{r}
filter(litters_df, !(gd_of_birth==20))

#this gives all the pups' gestational date of birth that doesn't equal to 20, !() flips it

#or you can do this:
filter(litters_df, gd_of_birth !=20)
```
filter two things at the same time
```{r}
filter(litters_df, gd0_weight >=22, gd_of_birth==20)
```
look at a categorical variable
```{r}
filter(litters_df, group == "Mod8")
```
filter two different categories, both Mod 8 and Con 7 use %in%
```{r}
filter(litters_df, group %in% c("Mod8", "Con7"))
```

##Mutate
create additional variables or modify exisitng variables
```{r}
mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight, #added wt_gain variable
  group = str_to_lower(group)) #modify the group variable by making everything lower case
```

##arrange
```{r}
arrange(litters_df, pups_born_alive)
# arrange by pups_born_alive variable

arrange(litters_df, pups_born_alive, gd0_weight)
```
## %>%
```{r}
litters_data_raw = read_csv("FAS_litters.csv")
litters_clean_name = janitor :: clean_names(litters_data_raw) # clean the names
litters_data_selected = select(litters_clean_name, -pups_survive) # take everything except for pups_survive
litters_mutated=mutate(litters_data_selected, wt_gain = gd18_weight - gd0_weight)
litters_without_missing = drop_na(litters_mutated, gd0_weight) # drop everything where gd0_weight is missing
```

USE THE PIPE OPERATOR INSTEAD

```{r}
litters_df = 
  read_csv("FAS_litters.csv") %>%
  janitor::clean_names() %>%
  select(-pups_survive) #select everything eexcept this variable
  mutate(wt_gain = gd18_weight - gd0_weight) %>%
  drop_na(gd0_weight)
  
```

## Tidy data
## pivot_longer
load the PULSE data

```{r}
pulse_data = 
  haven::read_sas("./public_pulse_data.sas7bdat") %>%
# use the haven package
  janitor::clean_names()

```

wide format to long format
```{r}
pulse_data_tidy=
  pulse_data %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit", 
    #the new column where all bdi_score go into "visit"
    names_prefix = "bdi_score_", 
    #get rid of alll the prefixes
    values_to= "bdi"   #the new column called bdi
  )
```

rewrite, combine, and extend (to add a mutate step)
```{r}
pulse_data = 
  haven::read_sas("./public_pulse_data.sas7bdat") %>%
# use the haven package
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  )%>%
  relocate(id, visit) %>%
  mutate(visit = recode(visit, "bl"="00m"))
# recode allows you to change observations, recode visit where bl is replaced by 00m
```






















